---
title: "Initial Korean covid 19"
author: "Kevin Ortega"
date: "October 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   


# Methods

$H_0$ :  There is no difference in dying depending on Sex.  

$H_A$ : There is a difference in dying depending on Sex.  

The data is from korea in 2022.  

result~sex  
this is catagorical ~ catagorical 
We will use helper file Cat_Cat
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  



```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```




The bar chart shows more males died than women.

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Among the people who died there were 64% of men who died and 36% of women who died.

## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P-value of Chi-squared test is 0.001 and the Fishcer exact test has a p-value of .00009. The odd of the fisher test also tastes that men are 2.3 more acceptable of dying from covid19 therefore the P-values are lower than 0.05 we reject the null hypothesis.


## Conclusions

$H_0$ :  There is no difference in dying depending on Sex.  

$H_A$ : There is a difference in dying depending on Sex.  

The data is from korea in 2022.  

result~sex  
this is catagorical ~ catagorical 
We will use helper file Cat_Cat


The bar chart shows more males died than women.

Among the people who died there were 64% of men who died and 36% of women who died.

The P-value of Chi-squared test is 0.001 and the Fishcer exact test has a p-value of .00009. The odd of the fisher test also tastes that men are 2.3 more acceptable of dying from covid19 therefore the P-values are lower than 0.05 we reject the null hypothesis.
